---
title: "Coronavirus"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
    vertical_layout: fill
---

```{r setup, include=FALSE}
#------------------ Packages ------------------
library(tidyverse)
library(flexdashboard)
library(covid19)
library(plotly)
library(DT)

covid19_df <- cvd_data()

coronavirus <-
  covid19_df %>%
  rename(
    Country.Region = `Country/Region`,
    Province.State = `Province/State`
  ) %>%
  pivot_longer(confirmed:deaths, names_to = "type", values_to = "cases")


#------------------ Parameters ------------------
# Set colors
# https://www.w3.org/TR/css-color-3/#svg-color
confirmed_color <- "purple"
active_color <- "#1f77b4"
recovered_color <- "forestgreen"
death_color <- "red"
#------------------ Data ------------------
df <- coronavirus %>% 
  filter(date == max(date)) %>%
  group_by(Country.Region, type) %>%
  summarise(total = sum(cases)) %>%
  pivot_wider(names_from =  type, 
              values_from = total) %>%
  mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(deaths), 0, deaths)) %>%
  arrange(-confirmed) %>%
  ungroup() %>%
  mutate(country = if_else(Country.Region == "United Arab Emirates", "UAE", Country.Region)) %>%
  mutate(country = if_else(country == "Mainland China", "China", country)) %>%
  mutate(country = if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  mutate(country = trimws(country)) %>%
  mutate(country = factor(country, levels = country))

df_daily <- coronavirus %>% 
  group_by(date, type) %>%
  summarise(total = sum(cases, na.rm = TRUE)) %>%
  pivot_wider(names_from = type,
              values_from = total) %>%
  arrange(date) %>%
  ungroup() %>%
  mutate(active =  confirmed - deaths - recovered) #%>%
# mutate(confirmed_cum = cumsum(confirmed),
#        deaths_cum = cumsum(deaths),
#        recovered_cum = cumsum(recovered),
#        active_cum = cumsum(active))


df1 <- coronavirus %>% filter(date == max(date))

```

```{r}
canada <- 
  coronavirus %>%
  filter(Country.Region == "Canada")  %>% 
  filter(date == max(date)) %>%
  group_by(Province.State, type) %>%
  summarise(total = sum(cases)) %>%
  pivot_wider(names_from =  type, 
              values_from = total) %>%
  mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(deaths), 0, deaths)) %>%
  arrange(-confirmed) %>%
  ungroup() %>%
  mutate(Province.State = factor(Province.State, levels = Province.State))

canada_daily <-
  coronavirus %>%
  filter(Country.Region == "Canada") %>%
  group_by(date, type) %>%
  summarise(total = sum(cases, na.rm = TRUE)) %>%
  pivot_wider(names_from = type,
              values_from = total) %>%
  arrange(date) %>%
  ungroup() %>%
  mutate(active =  confirmed - deaths - recovered)
  
```

# Canada Summary

## Row

```{r}
canada <- 
  coronavirus %>%
  filter(Country.Region == "Canada")  %>% 
  filter(date == max(date)) %>%
  group_by(Province.State, type) %>%
  summarise(total = sum(cases)) %>%
  pivot_wider(names_from =  type, 
              values_from = total) %>%
  mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(deaths), 0, deaths)) %>%
  arrange(-confirmed) %>%
  ungroup() %>%
  mutate(Province.State = factor(Province.State, levels = Province.State))
```


### confirmed {.value-box}

```{r}

valueBox(value = paste(format(sum(canada$confirmed), big.mark = ","), "", sep = " "), 
         caption = "Total Confirmed Cases", 
         icon = "fas fa-user-md", 
         color = confirmed_color)
```


### active {.value-box}

```{r}
valueBox(value = paste(format(sum(canada$unrecovered, na.rm = TRUE), big.mark = ","), " (",
                       round(100 * sum(canada$unrecovered, na.rm = TRUE) / sum(canada$confirmed), 1), 
                       "%)", sep = ""), 
         caption = "Active Cases", icon = "fas fa-ambulance", 
         color = active_color)
```

### recovered {.value-box}

```{r}
valueBox(value = paste(format(sum(canada$recovered, na.rm = TRUE), big.mark = ","), " (",
                       round(100 * sum(canada$recovered, na.rm = TRUE) / sum(canada$confirmed), 1), 
                       "%)", sep = ""), 
         caption = "Recovered Cases", icon = "fas fa-heartbeat", 
         color = recovered_color)
```

### deaths {.value-box}

```{r}

valueBox(value = paste(format(sum(canada$deaths, na.rm = TRUE), big.mark = ","), " (",
                       round(100 * sum(canada$deaths, na.rm = TRUE) / sum(canada$confirmed), 1), 
                       "%)", sep = ""),
         caption = "Death Cases", 
         icon = "fas fa-heart-broken", 
         color = death_color)
```


## Row

### Cases Distribution by Type

```{r}

canada %>%
  plot_ly( 
    x = ~ Province.State, 
    y = ~ unrecovered, 
    # text =  ~ confirmed, 
    # textposition = 'auto',
    type = "bar", 
    name = "Active",
    marker = list(color = active_color)
  ) %>%
  add_trace(
    y = ~ recovered, 
    # text =  ~ recovered, 
    # textposition = 'auto',
    name = "Recovered",
    marker = list(color = recovered_color)
  ) %>%
  add_trace(
    y = ~ deaths, 
    # text =  ~ deaths, 
    # textposition = 'auto',
    name = "Death",
    marker = list(color = death_color)
  ) %>%
  layout(
    barmode = 'stack',
    yaxis = list(title = "Total Cases"),
    xaxis = list(title = ""),
    hovermode = "compare",
    margin =  list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )

```

## Row {data-width=400}

### Cumulative Canadian Cases by Type

```{r}

# plot_ly(df_daily, x = ~date, y = ~active_cum, name = 'Active', type = 'scatter', mode = 'none', stackgroup = 'one', fillcolor = "#1f77b4") %>%
# add_trace(y = ~recovered_cum, name = 'Recovered', fillcolor = "green") %>%
# add_trace(y = ~deaths_cum, name = "Death", fillcolor = "red") %>%
#   layout(title = "",
#          xaxis = list(title = "",
#                       showgrid = FALSE),
#          yaxis = list(title = "Cumulative Number of Cases",
#                       showgrid = FALSE),
#          legend = list(x = 0.1, y = 0.9),
#                  hovermode = "compare")


canada_daily %>%
  plot_ly() %>%
    add_trace(
    x = ~ date,
    y = ~ confirmed,
    type = "scatter",
    mode = "lines+markers",
    name = "Confirmed",
    line = list(color = confirmed_color),
    marker = list(color = confirmed_color)
  ) %>%
  add_trace(
    x = ~ date,
    y = ~ active,
    type = "scatter",
    mode = "lines+markers",
    name = "Active",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  add_trace(
    x = ~ date,
    y = ~ recovered,
    type = "scatter",
    mode = "lines+markers",
    name = "Recovered",
    line = list(color = recovered_color),
    marker = list(color = recovered_color)) %>%
  add_trace(
    x = ~ date,
    y = ~ deaths,
    type = "scatter",
    mode = 'lines+markers',
    name = "Death",
    line = list(color = death_color),
    marker = list(color = death_color)
  ) %>%
  layout(
    title = "",
    yaxis = list(title = "Cumulative Total Cases"),
    xaxis = list(title = "Date"),
    legend = list(x = 0.1, y = 0.9),
    hovermode = "compare"
  )


```


### Recovery and Death Rates by Province

```{r}
df_summary <-
  canada %>%
  arrange(- confirmed) %>%
  select(province = Province.State, confirmed, recovered, deaths) %>%
  mutate(
    recover_rate = recovered / confirmed,
    death_rate = deaths / confirmed
  )  

df_summary %>%
  datatable(rownames = FALSE,
            colnames = c("Province", "Confirmed", "Recovered", "Death", "Recovery Rate", "Death Rate"),
            options = list(pageLength = nrow(df_summary), dom = 'tip')) %>%
  formatPercentage("recover_rate", 2) %>%
  formatPercentage("death_rate", 2) 
```

# Global Summary

## Row


### confirmed {.value-box}

```{r}

valueBox(value = paste(format(sum(df$confirmed), big.mark = ","), "", sep = " "), 
         caption = "Total Confirmed Cases", 
         icon = "fas fa-user-md", 
         color = confirmed_color)
```


### active {.value-box}

```{r}
valueBox(value = paste(format(sum(df$unrecovered, na.rm = TRUE), big.mark = ","), " (",
                       round(100 * sum(df$unrecovered, na.rm = TRUE) / sum(df$confirmed), 1), 
                       "%)", sep = ""), 
         caption = "Active Cases", icon = "fas fa-ambulance", 
         color = active_color)
```

### recovered {.value-box}

```{r}
valueBox(value = paste(format(sum(df$recovered, na.rm = TRUE), big.mark = ","), " (",
                       round(100 * sum(df$recovered, na.rm = TRUE) / sum(df$confirmed), 1), 
                       "%)", sep = ""), 
         caption = "Recovered Cases", icon = "fas fa-heartbeat", 
         color = recovered_color)
```

### deaths {.value-box}

```{r}

valueBox(value = paste(format(sum(df$deaths, na.rm = TRUE), big.mark = ","), " (",
                       round(100 * sum(df$deaths, na.rm = TRUE) / sum(df$confirmed), 1), 
                       "%)", sep = ""),
         caption = "Death Cases", 
         icon = "fas fa-heart-broken", 
         color = death_color)
```


## Row

### Cases Distribution by Type (Top 25 Countries)

```{r}

df %>%
  top_n(25, confirmed) %>%
  plot_ly( 
    x = ~ country, 
    y = ~ unrecovered, 
    # text =  ~ confirmed, 
    # textposition = 'auto',
    type = "bar", 
    name = "Active",
    marker = list(color = active_color)
  ) %>%
  add_trace(
    y = ~ recovered, 
    # text =  ~ recovered, 
    # textposition = 'auto',
    name = "Recovered",
    marker = list(color = recovered_color)
  ) %>%
  add_trace(
    y = ~ deaths, 
    # text =  ~ deaths, 
    # textposition = 'auto',
    name = "Death",
    marker = list(color = death_color)
  ) %>%
  layout(
    barmode = 'stack',
    yaxis = list(title = "Total Cases (log scale)", type = "log"),
    xaxis = list(title = ""),
    hovermode = "compare",
    margin =  list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )

```

## Row {data-width=400}

### Global Cumulative Cases by Type

```{r}

# plot_ly(df_daily, x = ~date, y = ~active_cum, name = 'Active', type = 'scatter', mode = 'none', stackgroup = 'one', fillcolor = "#1f77b4") %>%
# add_trace(y = ~recovered_cum, name = 'Recovered', fillcolor = "green") %>%
# add_trace(y = ~deaths_cum, name = "Death", fillcolor = "red") %>%
#   layout(title = "",
#          xaxis = list(title = "",
#                       showgrid = FALSE),
#          yaxis = list(title = "Cumulative Number of Cases",
#                       showgrid = FALSE),
#          legend = list(x = 0.1, y = 0.9),
#                  hovermode = "compare")


df_daily %>%
  plot_ly() %>%
    add_trace(
    x = ~ date,
    y = ~ confirmed,
    type = "scatter",
    mode = "lines+markers",
    name = "Confirmed",
    line = list(color = confirmed_color),
    marker = list(color = confirmed_color)
  ) %>%
  add_trace(
    x = ~ date,
    y = ~ active,
    type = "scatter",
    mode = "lines+markers",
    name = "Active",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  add_trace(
    x = ~ date,
    y = ~ recovered,
    type = "scatter",
    mode = "lines+markers",
    name = "Recovered",
    line = list(color = recovered_color),
    marker = list(color = recovered_color)) %>%
  add_trace(
    x = ~ date,
    y = ~ deaths,
    type = "scatter",
    mode = 'lines+markers',
    name = "Death",
    line = list(color = death_color),
    marker = list(color = death_color)
  ) %>%
  layout(
    title = "",
    yaxis = list(title = "Cumulative Total Cases"),
    xaxis = list(title = "Date"),
    legend = list(x = 0.1, y = 0.9),
    hovermode = "compare"
  )


```


### Recovery and Death Rates by Country

```{r}
df_summary <-
  df %>%
  arrange(- confirmed) %>%
  filter(confirmed >= 25) %>%
  select(country = Country.Region, confirmed, recovered, deaths) %>%
  mutate(
    recover_rate = recovered / confirmed,
    death_rate = deaths / confirmed
  )  

df_summary %>%
  datatable(rownames = FALSE,
            colnames = c("Country", "Confirmed", "Recovered", "Death", "Recovery Rate", "Death Rate"),
            options = list(pageLength = nrow(df_summary), dom = 'tip')) %>%
  formatPercentage("recover_rate", 2) %>%
  formatPercentage("death_rate", 2) 
```


# Global Trends

## Column {data-width=400}

```{r}

latest_date <- max(coronavirus$date)
date_before <- latest_date - 1

changes <-
  coronavirus %>%
  left_join(
    tibble(
      date = c(latest_date, date_before),
      is = c("today", "yesterday")
    )
  ) %>%
  filter(!is.na(is)) %>%
  select(-date) %>%
  pivot_wider(names_from = "is", values_from = "cases") %>%
  group_by(Country.Region, type) %>%
  summarise(
    yesterday = sum(yesterday),
    today = sum(today),
    change = today - yesterday
  ) %>%
  ungroup() %>%
  filter(change != 0) %>%
  mutate(pct_change = 1 - (yesterday / today))
```

### New Cases - Top 15 Countries (`r  latest_date`)

```{r}
changes %>%
  filter(type == "confirmed") %>%
  filter(change > 0) %>%
  arrange(-change) %>%
  mutate(country = factor(Country.Region, levels = Country.Region)) %>%
  top_n(n = 15, wt = change) %>%
  plot_ly(
    x = ~ country,
    y = ~ change,
    text = ~ change,
    textposition = 'auto',
    type = "bar"
  ) %>%
  layout(
    yaxis = list(title = str_glue("Reported Since {date_before}")),
    xaxis = list(title = ""),
    margin =  list(
      l = 10,
      r = 10,
      b = 10,
      t = 10,
      pad = 2
    )
  )

```

### New Deaths - Top 15 Countries (`r  latest_date`)

```{r}
changes %>%
  filter(type == "deaths") %>%
  filter(change > 0) %>%
  arrange(-change) %>%
  mutate(country = factor(Country.Region, levels = Country.Region)) %>%
  top_n(n = 15, wt = change) %>%
  plot_ly(
    x = ~ country,
    y = ~ change,
    text = ~ change,
    textposition = 'auto',
    type = "bar",
    marker = list(color = death_color)
  ) %>%
  layout(
    yaxis = list(title = str_glue("Reported Since {date_before}")),
    xaxis = list(title = ""),
    margin =  list(
      l = 10,
      r = 10,
      b = 10,
      t = 10,
      pad = 2
    )
  )

```


## Column {data-width=600}

### New Recoveries - Top 15 Countries (`r  latest_date`)

```{r}
changes %>%
  filter(type == "recovered") %>%
  filter(change > 0) %>%
  arrange(-change) %>%
  mutate(country = factor(Country.Region, levels = Country.Region)) %>%
  top_n(n = 15, wt = change) %>%
  plot_ly(
    x = ~ country,
    y = ~ change,
    text = ~ change,
    textposition = 'auto',
    type = "bar",
    marker = list(color = recovered_color)
  ) %>%
  layout(
    yaxis = list(title = str_glue("Reported Since {date_before}")),
    xaxis = list(title = ""),
    margin =  list(
      l = 10,
      r = 10,
      b = 10,
      t = 10,
      pad = 2
    )
  )

```

### Recovery and Death Rates for Countries with at Least 100 Cases

```{r}
coronavirus %>% 
  # filter(Country.Region != "Others") %>%
  group_by(Country.Region, type) %>%
  summarise(total_cases = sum(cases)) %>%
  pivot_wider(names_from = type, values_from = total_cases) %>%
  arrange(- confirmed) %>%
  filter(confirmed >= 100) %>%
  mutate(recover_rate = recovered / confirmed,
         death_rate = deaths / confirmed) %>% 
  mutate(recover_rate = if_else(is.na(recover_rate), 0, recover_rate),
         death_rate = if_else(is.na(death_rate), 0, death_rate)) %>%
  ungroup() %>%
  mutate(confirmed_normal = as.numeric(confirmed) / max(as.numeric(confirmed))) %>%
  plot_ly(y = ~ round(100 * recover_rate, 1),
          x = ~ round(100 * death_rate, 1),
          size = ~  log(confirmed),
          sizes = c(5, 70),
          type = 'scatter', mode = 'markers',
          color = ~ Country.Region,
          marker = list(sizemode = 'diameter' , opacity = 0.5),
          hoverinfo = 'text',
          text = ~paste("</br>", Country.Region, 
                        "</br> Confirmed Cases: ", confirmed,
                        "</br> Recovery Rate: ", paste(round(100 * recover_rate, 1), "%", sep = ""),
                        "</br> Death Rate: ",  paste(round(100 * death_rate, 1), "%", sep = ""))
  ) %>%
  layout(yaxis = list(title = "Recovery Rate", ticksuffix = "%"),
         xaxis = list(title = "Death Rate", ticksuffix = "%", 
                      dtick = 1, 
                      tick0 = 0),
         hovermode = "compare")

```   


# Data

```{r}
coronavirus %>% 
  select(Date = date, Province = Province.State, Country = Country.Region, `Case Type` = type, `Number of Cases` = cases) %>%
  datatable(rownames = FALSE,
            options = list(searchHighlight = TRUE, 
                           pageLength = 20), filter = 'top')
```



# About

**The Coronavirus Dashboard**

This Coronavirus dashboard provides an overview of the 2019 Novel Coronavirus COVID-19 (2019-nCoV) epidemic. This dashboard is built with R using the Rmakrdown framework and can easily reproduce by others. The code behind the dashboard available [here](https://github.com/RamiKrispin/coronavirus_dashboard)

**Data**

The input data for this dashboard is the [coronavirus](https://github.com/RamiKrispin/coronavirus) R package (dev version). The data and dashboard is refreshed on a daily bases. The raw data pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [repository](https://github.com/RamiKrispin/coronavirus-csv)




**Packages**

* Dashboard interface - the [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) package. 
* Visualization - the [plotly](https://plot.ly/r/) package
* Data manipulation - [dplyr](https://dplyr.tidyverse.org/), and [tidyr](https://tidyr.tidyverse.org/)
* Tables - the [DT](https://rstudio.github.io/DT/) package

**Deployment and reproducibly**

The dashboard was deployed to Github docs. If you wish to deploy and/or modify the dashboard on your Github account, you can apply the following steps:

* Fork the dashboard [repository](https://github.com/RamiKrispin/coronavirus_dashboard), or
* Clone it and push it to your Github package
* Here some general guidance about deployment of flexdashboard on Github page - [link](https://github.com/pbatey/flexdashboard-example)

For any question or feedback, you can either open an [issue](https://github.com/RamiKrispin/coronavirus_dashboard/issues) or contact me on [Twitter](https://twitter.com/Rami_Krispin).


